---
title: "PSY 525 Final Project"
author: "Amanda Moeller"
date: "4/19/2020"
output: html_document
---

<!-- 1. I suggest adding labels to your code chunks. -->
<!-- That way it is easier to monitor the progress of your document as it knits. -->

<!-- 2. I'd like to have seen a description of the dataset, your motivations for visualizing it, and the goals of the visualizations. -->
<!-- This is by no means a definitive recommendation, but I have taken to starting all of my R Markdown HTML reports with a "Purpose" and "Background" or "Set-up" section. -->
<!-- Also, because it is so easy, I like to provide commentary about what I am doing between each chunk. -->
<!-- If I am visualizing data, I like to comment on what I am seeing. -->

<!-- 3. This may be just a stylistic preference, but I prefer shorter headers and longer text in the paragraphs below the headers. -->
<!-- That allows you to generate a table of contents automatically using the `toc: true` parameter in your YAML header. -->

<!-- 4. I'm really happy with the progress you've made this term. Keep at it! -->

# This document contains a reproducible workflow for visualizing changes in mobility across the globe during the time of Covid-19.

## The data:
https://www.google.com/covid19/mobility/


```{r, include=FALSE}
library(tidyr)
library(tidyverse)
library(lubridate)
library(dplyr)
library(knitr)
library(ggplot2)
library(gganimate)
library(gifski) # this helps gganimate to turn png into videos/gifs
library(png) # see above note
library(readr)
library(scales) # package for scale formatting
library(kableExtra) # this and magick help format table
library(magick)
```


```{r, echo=FALSE, warning=FALSE, include=FALSE}
GoogleCovidCategories <- c("Retail and Recreation",
                           "Mobility trends for places like restaurants, cafes, shopping centers, theme parks, museums, libraries, and movie theaters.", 
                           "Grocery and Pharmacy", 
                           "Mobility trends for places like grocery markets, food warehouses, farmers markets, specialty food shops, drug stores, and pharmacies.", 
                           "Parks", 
                           "Mobility trends for places like local parks, national parks, public beaches, marinas, dog parks, plazas, and public gardens.", 
                           "Transit Stations", 
                           "Mobility trends for places like public transport hubs such as subway, bus, and train stations.", 
                           "Workplaces", 
                           "Mobility trends for places of work.", 
                           "Residential", 
                           "Mobility trends for places of residence")

Categories_df <- data.frame(matrix(GoogleCovidCategories, 
                        ncol=2, 
                        byrow=TRUE), 
stringsAsFactors=TRUE)
Categories_df
```

```{r, echo=FALSE, warning=FALSE}
GoogleCovidCategories <- data.frame(
  Category=c("Retail and Recreation",
                            "Grocery and Pharmacy",
                            "Parks",
                            "Transit Stations",
                            "Workplaces",
                            "Residential"),
  Category_Description =c(
                           "Mobility trends for places like restaurants, cafes, shopping centers, theme parks, museums, libraries, and movie theaters.", 
                         "Mobility trends for places like grocery markets, food warehouses, farmers markets, specialty food shops, drug stores, and pharmacies.", 
              "Mobility trends for places like local parks, national parks, public beaches, marinas, dog parks, plazas, and public gardens.", 
                          "Mobility trends for places like public transport hubs such as subway, bus, and train stations.",        "Mobility trends for places of work.", "Mobility trends for places of residence"))
  
  print(GoogleCovidCategories)
kable(head(GoogleCovidCategories[, 1:2]), "pandoc")
```


## The code used to generate plots for the percent change in mobility in Pennsylvania, USA is below.

#### Instructions for generating these plots for any U.S. state or specific county are also included. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Step 1: Load packages
```{r, warning=FALSE}
library(tidyr)
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
library(gganimate)
library(gifski) # this helps gganimate to turn png into videos/gifs
library(png) # see above note
library(readr)
library(scales) # package for scale formatting
```


### Step 2: Load the Community Mobility Covid-19 dataset    
```{r, warning=FALSE}
# Data is located here: https://www.google.com/covid19/mobility/
Global_Mobility_Report <- read.csv("https://www.gstatic.com/covid19/mobility/Global_Mobility_Report.csv", head=TRUE)
head(Global_Mobility_Report) %>% as_tibble()
```

### Step 3: Here, I create a subset of the Global Report to include only location data from Pennsylvania.

#### To obtain how mobility to various places has changed within different states across the USA, simply replace the following:

#### -- "sub_region_1" == 'Pennsylvania' with the name of another state 
#### AND/OR
#### -- "sub_region_2" == 'COUNTY NAME' if you want county-level data only
```{r}
# Create a new data frame with only Pennsylvania mobility data
PA_community_mobility <- subset(Global_Mobility_Report, 
                                sub_region_1 == 'Pennsylvania', 
                                select = c("sub_region_1", 
                                           "sub_region_2", 
                                           "date",
                                    "retail_and_recreation_percent_change_from_baseline",
                                           "grocery_and_pharmacy_percent_change_from_baseline",
                                           "parks_percent_change_from_baseline",
                                           "transit_stations_percent_change_from_baseline",
                                           "workplaces_percent_change_from_baseline",
                                           "residential_percent_change_from_baseline"))

# Change date column to as.date so R can read it as a date
PA_community_mobility$date <- as.Date(PA_community_mobility$date)
```

## Step 4: Plot out the mobility change for each place category between the time period of February 15, 2020 to April 10, 2020:


### Retail and recreation change in Pennsylvania from 02/15/2020 to 04/10/2020
STATIC PLOT
```{r, echo=TRUE, warning=FALSE}
# Static retail and recreation mobility plot:
staticPlot_RetailRecreation_mobilitychange <- 
  ggplot(data=PA_community_mobility)+
  geom_point(mapping=aes(x=date, 
                         y=retail_and_recreation_percent_change_from_baseline, 
                         color=retail_and_recreation_percent_change_from_baseline))+
  theme_minimal() +
  scale_color_gradient(low = "#0091ff", 
                       high = "#f0650e")  +
  theme_linedraw() 
# Display the plot:
print(staticPlot_RetailRecreation_mobilitychange) + labs(x="Date", y="Retail and Recreation Mobility Change in PA") 
```


### Retail and recreation mobility change in Pennsylvania from 02/15/2020 to 04/10/2020 (Animated plot)
ANIMATED PLOT
```{r, echo=TRUE, warning=FALSE}
# Animated retail and recreation mobility plot:
animPlot_RetailRecreation_mobilitychange <- 
  ggplot(PA_community_mobility,
         aes(x=date, 
             y=retail_and_recreation_percent_change_from_baseline)) +
  geom_point(aes(color=retail_and_recreation_percent_change_from_baseline))

animPlot_RetailRecreation_mobilitychange <- 
  animPlot_RetailRecreation_mobilitychange+
  transition_time(PA_community_mobility$date) +
  shadow_mark() + 
  scale_x_date(labels=date_format("%m/%d")) +
  xlab("Date") + 
  ylab("Retail and Recreation Mobility Change in PA")+
  scale_color_gradient(low = "#0091ff", 
                       high = "#f0650e")+
  theme(legend.position="none") +
  theme_linedraw() +
  theme(legend.position="none")

# Display the plot:
animated_RetailRecreation_plot <- animate(animPlot_RetailRecreation_mobilitychange, width=700, height=500)
animated_RetailRecreation_plot
```

## Grocery and Pharmacy 
### Mobility change in Pennsylvania 
Static plot
```{r, warning=FALSE}
# Static grocery and pharmacy mobility plot:
staticPlot_GroceryPharmacy_mobilitychange <- 
  ggplot(data=PA_community_mobility)+
  geom_point(mapping=aes(x=date, 
                         y=grocery_and_pharmacy_percent_change_from_baseline, 
                         color=grocery_and_pharmacy_percent_change_from_baseline))+
  theme_minimal() +
  scale_color_gradient(low = "#0091ff", 
                       high = "#f0650e")  +
  theme_linedraw() 
# Display the plot:
print(staticPlot_GroceryPharmacy_mobilitychange) + labs(x="Date", y="Grocery and Pharmacy Mobility Change in PA") 
```
ANIMATED PLOT
```{r, warning=FALSE}
# Animated grocery and pharmacy mobility plot:
animPlot_GroceryPharmacy_mobilitychange <- 
  ggplot(PA_community_mobility,
         aes(x=date, 
             y=grocery_and_pharmacy_percent_change_from_baseline)) +
  geom_point(aes(color=grocery_and_pharmacy_percent_change_from_baseline))

animPlot_GroceryPharmacy_mobilitychange <- 
  animPlot_GroceryPharmacy_mobilitychange+
  transition_time(PA_community_mobility$date) +
  shadow_mark() + 
  scale_x_date(labels=date_format("%m/%d")) +
  xlab("Date") + 
  ylab("Grocery and Pharmacy Mobility Change in PA")+
  scale_color_gradient(low = "#0091ff", 
                       high = "#f0650e")+
  theme(legend.position="none") +
  theme_linedraw() +
  theme(legend.position="none")

# Display the plot:
animated_GroceryPharmacy_plot <- animate(animPlot_GroceryPharmacy_mobilitychange, width=700, height=500)
animated_GroceryPharmacy_plot
```

## Parks 
### Mobility change in Pennsylvania
Static plot
```{r, warning=FALSE}
# Static parks mobility plot:
staticPlot_Parks_mobilitychange <- 
  ggplot(data=PA_community_mobility)+
  geom_point(mapping=aes(x=date, 
                         y=parks_percent_change_from_baseline, 
                         color=parks_percent_change_from_baseline))+
  theme_minimal() +
  scale_color_gradient(low = "#0091ff", 
                       high = "#f0650e") +
  theme_linedraw() 
  
# Display the plot:
print(staticPlot_Parks_mobilitychange) + labs(x="Date", y="Parks Mobility Change in PA") 

```

Animated plot
```{r, warning=FALSE}
# Animated parks mobility plot:
animPlot_Parks_mobilitychange <- 
  ggplot(PA_community_mobility,
         aes(x=date, 
             y=parks_percent_change_from_baseline)) +
  geom_point(aes(color=parks_percent_change_from_baseline))

animPlot_Parks_mobilitychange <- 
  animPlot_Parks_mobilitychange+
  transition_time(PA_community_mobility$date) +
  shadow_mark() + 
  scale_x_date(labels=date_format("%m/%d")) +
  xlab("Date") + 
  ylab("Parks Mobility Change in PA")+
  scale_color_gradient(low = "#0091ff", 
                       high = "#f0650e")+
  theme(legend.position="none")

# Display the plot:
animated_Parks_plot <- animate(animPlot_Parks_mobilitychange, width=700, height=500)
animated_Parks_plot
```


## Transit Stations 
### Mobility change in Pennsylvania 
Static plot
```{r, warning=FALSE}
# Static transit stations mobility plot:
staticPlot_TransitStations_mobilitychange <- 
  ggplot(data=PA_community_mobility)+
  geom_point(mapping=aes(x=date, 
                         y=transit_stations_percent_change_from_baseline, 
                         color=transit_stations_percent_change_from_baseline))+
  theme_minimal() +
  scale_color_gradient(low = "#0091ff", 
                       high = "#f0650e")  +
  theme_linedraw() 
# Display the plot:
print(staticPlot_TransitStations_mobilitychange) + labs(x="Date", y="Transit Stations Mobility Change in PA") 

```

Animated plot
```{r, warning=FALSE}
# Animated transit stations mobility plot:
animPlot_TransitStations_mobilitychange <- 
  ggplot(PA_community_mobility,
         aes(x=date, 
             y=transit_stations_percent_change_from_baseline)) +
  geom_point(aes(color=transit_stations_percent_change_from_baseline))

animPlot_TransitStations_mobilitychange <- 
  animPlot_TransitStations_mobilitychange+
  transition_time(PA_community_mobility$date) +
  shadow_mark() + 
  scale_x_date(labels=date_format("%m/%d")) +
  xlab("Date") + 
  ylab("Transit Stations Mobility Change in PA")+
  scale_color_gradient(low = "#0091ff", 
                       high = "#f0650e")+
  theme(legend.position="none") +
  theme_linedraw() 

# Display the plot:
animated_TransitStations_plot <- animate(animPlot_TransitStations_mobilitychange, 
                                         width=700, 
                                         height=500)
animated_TransitStations_plot
```


## Workplaces 
### Mobility change in Pennsylvania from 02/15/2020 to 04/10/2020 
Static plot
```{r, warning=FALSE}
# Static workplace mobility plot:
staticPlot_Workplaces_mobilitychange <- 
  ggplot(data=PA_community_mobility)+
  geom_point(mapping=aes(x=date, 
                         y=workplaces_percent_change_from_baseline, 
                         group=workplaces_percent_change_from_baseline, 
                         color=workplaces_percent_change_from_baseline))+
  theme_minimal() +
  scale_color_gradient(low = "#0091ff", 
                       high = "#f0650e") +
  theme_linedraw() 
# Display the plot:
print(staticPlot_Workplaces_mobilitychange) + labs(x="Date", y="Workplaces Mobility Change in PA") 

```

Animated plot
```{r, warning=FALSE}
# Animated workplace mobility plot:
animPlot_Workplaces_mobilitychange <- 
  ggplot(PA_community_mobility,
         aes(x=date, 
             y=workplaces_percent_change_from_baseline)) +
  geom_point(aes(color=workplaces_percent_change_from_baseline))

animPlot_Workplaces_mobilitychange <- 
  animPlot_Workplaces_mobilitychange+
  transition_time(PA_community_mobility$date) +
  shadow_mark() + 
  scale_x_date(labels=date_format("%m/%d")) +
  xlab("Date") + 
  ylab("Workplace Mobility Change in PA")+
  scale_color_gradient(low = "#0091ff", 
                       high = "#f0650e")+
  theme(legend.position="none")+
  theme_linedraw() 

# Display the plot:
animated_Workplaces_plot <- animate(animPlot_Workplaces_mobilitychange, width=700, height=500)
animated_Workplaces_plot
```


## Residential Area 
### Mobility change in Pennsylvania from 02/15/2020 to 04/10/2020
Static plot
```{r, warning=FALSE}
# Residential area mobility plot:
staticPlot_Residential_mobilitychange <- 
  ggplot(data=PA_community_mobility)+
  geom_point(mapping=aes(x=date, 
                         y=residential_percent_change_from_baseline, 
                         color=residential_percent_change_from_baseline))+
  theme_minimal() +
  scale_color_gradient(low = "#0091ff", 
                       high = "#f0650e")  +
  theme_linedraw() 

# Display the plot:
print(staticPlot_Residential_mobilitychange) + labs(x="Date", y="Residential Area Mobility Change in PA")
```

Residential Area mobility change in Pennsylvania from 02/15/2020 to 04/10/2020 (Animated plot)
```{r, warning=FALSE}
# Animated residential mobility plot:
animPlot_Residential_mobilitychange <- 
  ggplot(PA_community_mobility,
         aes(x=date, 
             y=residential_percent_change_from_baseline)) +
  geom_point(aes(color=residential_percent_change_from_baseline))

animPlot_Residential_mobilitychange <- 
  animPlot_Residential_mobilitychange+
  transition_time(PA_community_mobility$date) +
  shadow_mark() + 
  scale_x_date(labels=date_format("%m/%d")) +
  xlab("Date") + 
  ylab("Residential Area Mobility Change in PA")+
  scale_color_gradient(low = "#0091ff", 
                       high = "#f0650e")+
  theme(legend.position="none") +
  theme_linedraw() 

# Display the plot:
animated_Residential_plot <- animate(animPlot_Residential_mobilitychange, 
                                         width=700, 
                                         height=500)
animated_Residential_plot
```









